---
title: "Medical_No_Show_Data_EDA"
author: "Team No Show (Chhavi Sharma, Prahasan Gadugu and Supriya Ayalur Balasubramanian)"
date: "April 5, 2019"
output: html_document
---


####Citation:
https://medium.com/@williamkoehrsen/exploratory-data-analysis-with-r-f9d3a4eb6b16

https://www.kaggle.com/yousuf28/medical-appointment-no-show-in-r/report

https://www.kaggle.com/jph84562/data-exploration-and-visualization/report

https://www.kaggle.com/cnjn22/exploratory-analysis-for-medical-appointment

https://www.kaggle.com/skirmer/excavating-insights-in-medical-no-shows


####Dataset Source:
https://www.kaggle.com/joniarroba/noshowappointments/version/3

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

# Load the required packages.
suppressMessages(library(ggpubr))
suppressMessages(library(ggplot2))
suppressMessages(library(plyr))
suppressMessages(library(dplyr))
suppressMessages(library(reshape2))
suppressMessages(library(Hmisc))
suppressMessages(library(GGally))
suppressMessages(library(lubridate))
suppressMessages(library(stringr))
suppressMessages(library(scales))
suppressMessages(library(corrplot))
suppressMessages(library(GGally))
suppressMessages(library(gridExtra))
suppressMessages(library(grid))
```



## Reading the data

```{r}
no_show_df <- read.csv('No-show-Issue-Comma-300k.csv',header = TRUE)
str(no_show_df)
```



## Data Preparation and Cleaning Phase:

```{r}

summary(no_show_df)
#Age has got min as -2 and max as 113 need to go in depth, removing the negative ages.

no_show_df <- filter(no_show_df, Age >= 0)

#Awaiting time has been given which is further converted to waiting time by changing it into positive values.
no_show_df$AwaitingTime = -(no_show_df$AwaitingTime)

#Status variable is converted into Status _int with numeric labels 0 (for no show), 1(show up)
no_show_df$Status_int <- factor(no_show_df$Status, levels=c("Show-Up", "No-Show"),labels = c(0, 1))


#Appointment registration has been converted to ymd_hms format to extract the hour information.
no_show_df$Status_int <- as.numeric(no_show_df$Status_int)
no_show_df$AppointmentRegistration_ymd_hms = ymd_hms(no_show_df$AppointmentRegistration)
no_show_df$ApointmentData <-gsub(pattern = "[A-Z]", replacement = " ",
                                x = no_show_df$ApointmentData)
no_show_df$AppointmentRegistration <- 
  gsub(pattern = "[A-Z]", replacement = " ", 
       x = no_show_df$AppointmentRegistration)
no_show_df$ApointmentData <- as.Date(no_show_df$ApointmentData)
no_show_df$AppointmentRegistration <- 
  as.Date(no_show_df$AppointmentRegistration)


#Appointment Date column has been modified by dividing it into year, month and day.
no_show_df <-transform(no_show_df, month = format(ApointmentData, "%m"))
no_show_df <- transform(no_show_df, year = format(ApointmentData, "%Y"))
no_show_df <- transform(no_show_df, day = format(ApointmentData, "%d"))
no_show_df$year <- as.integer(as.character(no_show_df$year))
no_show_df$month <- as.integer(as.character(no_show_df$month))
no_show_df$day <- as.integer(as.character(no_show_df$day))

str(no_show_df)
```

#Correlation Matrix

```{r fig.height= 7, fig.width= 5}
mydata.cor = cor(no_show_df[,c("Status_int","Age", "Diabetes", "Alcoolism","Handcap",
                             "HiperTension" , "Smokes", "Scholarship", 
                             "Tuberculosis", "AwaitingTime", "month", "year", 
                             "day","Sms_Reminder")])
colnames(mydata.cor) <- c("STATUS","Age", "Diabetes", "Alcoholism","Handicap",
                             "HyperTension" , "Smokes", "Scholarship", 
                             "Tuberculosis", "Awaiting.Time", "Appointment.Month", "Appointment.Year", 
                             "Appointment.Day","SMS.Reminder")
gg_mat = ggcorr(mydata.cor,label=TRUE,hjust = 0.9, size = 3, color = "black", layout.exp = 2)+ labs(title = "Correlation Matrix", subtitle = "Determination of Correlation of Factors w.r.t \"Status\" variable")
gg_mat
```


##Exploratory Data Analysis starts here:

#Plot 1:
#Absence rate vs Age:


```{r}

#Age behaviour

no_shows_age_lessthan_90 = filter(no_show_df, Age<90)
no_shows_by_age<-group_by(no_shows_age_lessthan_90, Age)
no_shows_by_age <- dplyr::summarize(no_shows_by_age,
                             absence_rate = mean(Status_int),
                             n = n())

#Scatter plot and trend
ggplot(aes(x = Age, y = absence_rate-1), data = no_shows_by_age) + geom_smooth(method = "loess")+
  geom_point(size = 1) +  labs(x = 'Age (years)', y = 'No-Show Rate %' , 
                               title = 'No-show rate vs Age of Patient') +   scale_x_continuous(breaks=seq(0,100,5))


#creating bins of Age and plotting the Absence rate vs Age bins:


no_shows_by_age$age_bucket <- cut(no_shows_by_age$Age, 
                                  breaks= seq(0,90, 5), include.lowest=TRUE)
no_shows_by_age_bucket <- group_by(no_shows_by_age, age_bucket)
no_shows_by_age_bucket <- dplyr::summarize(no_shows_by_age_bucket,
                                           absence_rate = mean(absence_rate-1),
                                           n = n())

gg_age = ggplot(aes(x = age_bucket, y = 100*absence_rate), 
       data= no_shows_by_age_bucket) + 
  geom_bar(stat='identity', color='black', 
           fill = 'orange', lwd = 0.8, width=1) + 
  theme(axis.text.x=element_text(angle = 60, hjust=1)) + 
  labs(x = 'Age Group', y = 'No-Show Rate in %', 
       title = 'No-Show Rate vs Age Group') + 
  scale_y_continuous(breaks=seq(-40, 40, 10))
gg_age

```

#Plot 2:
#No-show rate vs Waiting Time bins

```{r}

#Waiting time behaviour
no_shows_by_wait <- group_by(no_show_df, AwaitingTime)
no_shows_by_wait <- dplyr::summarize(no_shows_by_wait, 
                                     absence_rate = mean(Status_int),
                                     n = n())


# No show rate vs age scatter plot.
ggplot(aes(x = AwaitingTime, y = 100 * (absence_rate-1)), data = no_shows_by_wait) + 
  geom_point(color='black') + geom_smooth(method = "loess")+
  labs(x = 'Wait (days)' , y = 'No-show Rate %', 
       title = 'No-show Rate vs Waiting Time')

#Filtering out the ages less than 90 days
no_shows_by_wait <- filter(no_shows_by_wait, AwaitingTime< 90)


#Cutting the wait time into bins of 10 days each like the age.

no_shows_by_wait$wait_bucket<-cut(no_shows_by_wait$AwaitingTime, 
                                  breaks = seq(0,90,10),include.lowest = TRUE)

no_shows_by_wait_bucket <- group_by(no_shows_by_wait, wait_bucket)
no_shows_by_wait_bucket <- dplyr::summarize(no_shows_by_wait_bucket,
                                           absence_rate = mean(absence_rate-1),
                                           n = n())


#Relative absence rate is calculated as the trend in the absolute absence rate is not significant.
no_shows_by_wait_bucket$relative_absence_rate <- 
  with(no_shows_by_wait_bucket, 
       100 *(absence_rate - mean(absence_rate)) / mean(absence_rate))


# Graph the relative absence rate for each age bucket
gg_wait = ggplot(aes(x = wait_bucket, y = relative_absence_rate), 
       data = no_shows_by_wait_bucket) + 
  geom_bar(stat='identity', color='black',fill='darkgreen',lwd = 0.8, width = 1) + 
  scale_y_continuous(breaks=seq(-20, 50, 5)) + 
  labs(x = 'Wait Time Group (days)', y = 'Relative No-Show Rate in %', 
       title = 'Relative No-Show Rate vs Waiting Time  Group')

gg_wait

```



#Plot 3:
#No show rate vs month

```{r fig.width = 7, fig.height = 5}

##Month plot 1:

no_shows_by_month <- group_by(no_show_df, month)
no_shows_by_month <- dplyr::summarize(no_shows_by_month,
                                      absence_rate = mean(Status_int),
                                      n = n())

month_labels = c('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun', 'Jul' ,'Aug', 'Sep', 
                 'Oct', 'Nov' , 'Dec')
no_shows_by_month$month_name <- month_labels
no_shows_by_month$month_name <- 
  factor(no_shows_by_month$month_name, 
         levels=unique(no_shows_by_month$month_name))
gg_month_line = ggplot(no_shows_by_month, aes(x = month_name, y = 100*(absence_rate-1))) + 
  geom_point(color = 'red', size = 1.8) +
  labs(x = 'Appointment Month', y = 'Absence Rate %', 
       title='Absence Rate vs Appointment Month')+ 
  geom_line(group =1, color = 'blue')
gg_month_line
no_shows_by_month <-
no_show_df %>% 
  group_by(month, Gender) %>% 
  dplyr::count(Status)

##Month plot 2

no_shows_by_month <- tidyr::spread(no_shows_by_month, key = Status, value = n)

colnames(no_shows_by_month) <- c("Month", "Gender", "Noshow", "Show")

no_shows_by_month <-
no_shows_by_month %>% 
  mutate(Noshowrate = round(Noshow / (Noshow + Show), digits = 3))

gg_month = no_shows_by_month %>%
  group_by(Month) %>% 
  summarise(Noshowrate = mean(Noshowrate)) %>%
  ggplot(aes(x = Month, y = 100*Noshowrate)) +
  geom_bar(stat = "identity", fill = "darkgreen") +
  scale_x_discrete(limits = c(1, 2, 3, 4, 5, 6, 7,8,9,10,11,12),
                   labels = c("Jan", "Feb", "Mar", "Apr", "May","Jun", "Jul", "Aug","Sep","Oct","Nov","Dec")) +
  xlab("Month of the year") +
  ylab("No-Show Rate in %") +
  ggtitle("No-Show Rate Against month in a year") 

gg_month


```




#Plot 4
#No show rate by week day:

```{r}
no_show_byweekday <-
no_show_df %>% 
  group_by(DayOfTheWeek, Gender) %>% 
  dplyr::count(Status)


#Taking the information from week day counting the Staus and assigning to 
no_show_byweekday <- tidyr::spread(no_show_byweekday, key = Status, value = n)

colnames(no_show_byweekday) <- c("DayOfTheWeek", "Gender", "Noshow", "Show")
no_show_byweekday <-
no_show_byweekday %>% 
  mutate(Noshowrate = round(Noshow / (Noshow + Show), digits = 3))
no_show_byweekday$Gender <- factor(no_show_byweekday$Gender, levels = c("M", "F"))
week_day = no_show_byweekday %>%
  group_by(DayOfTheWeek) %>% 
  summarise(Noshowrate = mean(Noshowrate))
gg_week_day = ggplot(week_day, aes(x = DayOfTheWeek, y = 100*Noshowrate))+
  geom_bar(stat = "identity", fill = "purple") +
  scale_x_discrete(limits = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"),
                   labels = c("Mon", "Tues", "Wed", "Thurs", "Fri", "Sat", "Sun")) +
  ggtitle("No-Show Rate Against Day of Week") + xlab("Day of the week")+ylab("No Show rate in %")
gg_week_day
  
no_show_df$DayOfTheWeek <- factor(no_show_df$DayOfTheWeek,levels=c("Monday","Tuesday","Wednesday",
                                                       "Thursday","Friday","Saturday","Sunday"))


gg_imbal_weekday = ggplot(no_show_df,aes(x=DayOfTheWeek,fill=DayOfTheWeek))+
  geom_bar(fill = "darkcyan")+xlab("Day of the week")+ylab("Count")+ggtitle("Count of Appointments vs Days of the week")

gg_week_day_prop = ggplot(no_show_df, aes(x=DayOfTheWeek,fill=Status))+
  geom_bar(stat="count",position="fill")+xlab("Day of the week")+ylab("Proportion of Show and No-Show")+ggtitle("Proportion of Show and No-show w.r.t days of the week")


gg_dow = grid.arrange(gg_imbal_weekday,gg_week_day_prop)
gg_dow
```

#Plot 5:
#Absence rate vs Appointment day

```{r fig.width = 10, fig.height = 6}
no_shows_by_day <- group_by(no_show_df, day)
no_shows_by_day <- dplyr::summarize(no_shows_by_day,
                                    absence_rate = mean(Status_int),
                                    n = n())
gg_appt_day = ggplot(aes(x = day, y = 100*(absence_rate-1)), data = no_shows_by_day) + 
  geom_point(color = 'red')  + labs(x = 'Day of Appointment', 
                                    y = 'No-Show Rate in %',
                                    title='No-Show Rate vs Appointment Day') +
  scale_x_continuous(breaks=seq(1,31,1)) + geom_line(color='blue')+theme(axis.text.x=element_text(angle = 60, hjust=1))


gg_appt_day
```


#Plot 6:
#No-show Rate vs No.of SMS Reminders
```{r}
no_shows_by_sms <- group_by(no_show_df, Sms_Reminder)
no_shows_by_sms <- dplyr::summarize(no_shows_by_sms, 
                                    absence_rate = mean(Status_int),
                                    n = n())
gg_sms = ggplot(no_shows_by_sms,aes(x = Sms_Reminder, y = 100*(absence_rate-1))) + 
  geom_bar(stat='identity', color='black', fill='peru') + 
  labs(x = 'Number of SMS Reminders', y = 'No-show Rate %' , 
       title = 'No-Show Rate vs  Number of \n SMS Reminders')+scale_y_continuous(breaks=seq(0, 40, 10))
gg_sms


gg_sms_prop = ggplot(no_show_df, aes(x=Sms_Reminder,fill=Status))+
  geom_bar(stat="count",position="fill")+xlab("Number of SMS Reminders")+ylab("Proportion of Show and No-Show")+ggtitle("Proportion of Show and No-show \n w.r.t no.of SMS Reminders")
gg_sms_prop


gg_sms_prop = grid.arrange(gg_sms,gg_sms_prop,nrow =1)
gg_sms_prop
```


#Plot 7:
#Absence rate vs Age by Text Reminders

```{r}
no_show_df_avg <- filter(no_show_df, Age<90)
ggplot(aes(x = Age, y = Sms_Reminder), data = no_show_df_avg) +
  geom_line(stat='summary', fun.y = mean, color='darkblue') + 
  labs(x = 'Age (years)', y = 'Number of SMS Reminders for each appointment', 
       title='Average SMS Reminders vs Age' ) + 
  scale_x_continuous(breaks= seq(0,90, 10))


no_show_df_sms_01 <- filter(no_show_df, Sms_Reminder<2)
no_shows_by_texts <- group_by(no_show_df_sms_01, Age)
no_shows_by_texts <- filter(no_shows_by_texts, Age < 90)
no_shows_by_texts$Sms_Reminder <- factor(no_shows_by_texts$Sms_Reminder)
gg_age_sms = ggplot(aes(x = Age, y = 100*(Status_int-1) ), data = no_shows_by_texts) + 
  geom_line(aes(color=Sms_Reminder), stat = 'summary', fun.y = mean, lwd = 1.2) + 
  labs(x = 'Age (yrs)', y = 'Absence Rate %', 
       title='Absence Rate vs Age by Text Reminders' ) + 
  scale_color_manual(name='Number of SMS Reminders', 
                     values=c('red','blue', 'slategray4'))

gg_age_sms
```


#Plot 8:
#No-show rate vs Reg_hour

```{r fig.width= 5, fig.height = 7}


no_show_by_hour <- group_by(no_show_df, RegistrationHour=hour(AppointmentRegistration_ymd_hms)) 

no_show_by_hour<- dplyr::summarize(no_show_by_hour,absence_rate = mean(Status_int),
                             n = n())
gg_reghour = ggplot(no_show_by_hour, aes(x=RegistrationHour, y=100*(absence_rate-1))) + geom_bar(stat="identity",fill="brown") + scale_fill_discrete("Registration Hour")+xlab("Registration Hour")+ylab("Absence Rate in %")+ggtitle("Absence Rate vs Registration hour in a day")

ggplot(no_show_df, aes(x=hour(AppointmentRegistration_ymd_hms),fill=Status))+
  geom_bar(stat="count",position="fill")+xlab("Registration Hour")+ylab("Proportion of Show and No-Show")+ggtitle("Proportion of Show and No-show \n w.r.t Registration hour in a day")

ggplot(no_show_df, aes(x=hour(AppointmentRegistration_ymd_hms),fill=hour(AppointmentRegistration_ymd_hms)))+
  geom_histogram(fill = "black", binwidth = 0.5)+xlab("Registration Hour")+ylab("Count of Appointments")+ggtitle("Count of Appointments \n w.r.t Registration hour in a day")
gg_reghour


no_show_df$appt_hr_bucket <- cut(hour(no_show_df$AppointmentRegistration_ymd_hms), 
                                  breaks= seq(0,30,5), include.lowest=TRUE)
no_shows_by_appt_hr_bucket <- group_by(no_show_df, appt_hr_bucket)
no_shows_by_appt_hr_bucket <- dplyr::summarize(no_shows_by_appt_hr_bucket,
                                           absence_rate = mean(Status_int),
                                           n = n())

gg_apptr_bins = ggplot(aes(x = appt_hr_bucket, y = 100*(absence_rate-1)), 
       data= no_shows_by_appt_hr_bucket) + 
  geom_bar(stat='identity', color='black', 
           fill = 'orange', lwd = 0.8, width=1) + 
  theme(axis.text.x=element_text(angle = 60, hjust=1)) + 
  labs(x = 'Registration Hour', y = 'No-Show Rate in %', 
       title = 'No-Show Rate vs Appointment Registration \n hour Group') 


gg_count_rehour = ggplot(no_show_df, aes(appt_hr_bucket))+
  geom_bar(fill = "purple")+xlab("Registration hour")+ylab("Count")+ggtitle("Count of appointments \n w.r.t Registration hour in a day")

gg_prop_rehour = ggplot(no_show_df, aes(x=appt_hr_bucket,fill=Status))+
  geom_bar(stat="count",position="fill")+xlab("Registration hour")+ylab("Proportion of \n Show and No-Show")+ggtitle("Proportion of Show and No-show \n w.r.t Registration hour in a day")


gg_apptr_bins
gg_count_rehour
gg_prop_rehour


grid.arrange(gg_count_rehour,gg_prop_rehour,ncol = 1)

```




#Plot 9
#Slide 9: Seasonal and holiday wise behaviour


```{r}
#This cell takes time please be patient:
#Seasonal Behaviour
#Particular Date

# First plot for 2014

no_show_df_2014 = no_show_df[which(no_show_df$year == 2014),]
no_show_df_2015 = no_show_df[which(no_show_df$year == 2015),]
plot1 <-  ggplot(no_show_df_2014, aes(x = ApointmentData, y = (100*Status_int)-100)) +
  geom_point(aes(color = Gender),stat = 'summary', fun.y = mean)+
  labs(x = 'Appt. Date (M-D)', y = 'Absence Rate %',
       title = 'Absence Rate vs. Day of Year 2014') + 
  scale_x_date(date_breaks = "1 month", date_labels = "%m-%d") + 
  theme(axis.text.x  =element_text(angle = 70, hjust = 1))+geom_vline(aes(xintercept=(as.numeric(as.Date("2014-12-25")))), 
             linetype = 2, color = 'brown') + 
  geom_text(aes(x=(as.Date("2014-12-25")), y= 60, label='Christmas'),
            size=3, angle=90, vjust=-0.4, hjust=0)+geom_vline(aes(xintercept=(as.numeric(as.Date("2014-08-01")))), 
             linetype = 2, color = 'brown') + 
  geom_text(aes(x=(as.Date("2014-08-01")), y= 60, label='Fall'),
            size=3, angle=90, vjust=-0.4, hjust=0)+geom_vline(aes(xintercept=(as.numeric(as.Date("2014-05-01")))), 
             linetype = 2, color = 'brown') + 
  geom_text(aes(x=(as.Date("2014-05-01")), y= 60, label='Summer'),
            size=3, angle=90, vjust=-0.4, hjust=0)+geom_vline(aes(xintercept=(as.numeric(as.Date("2014-03-01")))), 
             linetype = 2, color = 'brown') + 
  geom_text(aes(x=(as.Date("2014-03-01")), y= 60, label='Spring'),
            size=3, angle=90, vjust=-0.4, hjust=0)+geom_vline(aes(xintercept=(as.numeric(as.Date("2014-11-15")))), 
             linetype = 2, color = 'brown') + 
  geom_text(aes(x=(as.Date("2014-11-15")), y= 60, label='Thanksgiving and Winter'),
            size=3, angle=90, vjust=-0.4, hjust=0)
  
# Second plot for 2015
plot2 <- ggplot(no_show_df_2015, aes(x = ApointmentData, y = (100*Status_int)-100)) + 
  geom_point(aes(color = Gender),stat = 'summary', fun.y = mean)  + 
  labs(x = 'Appt. Date (M-D)', y = 'Absence Rate %', 
       title = 'Absence Rate vs. Day of Year 2015') +
  scale_x_date(date_breaks = "1 month", date_labels = "%m-%d") + 
  theme(axis.text.x  =element_text(angle = 70, hjust = 1))+geom_vline(aes(xintercept=(as.numeric(as.Date("2015-12-25")))), 
             linetype = 2, color = 'brown') + 
  geom_text(aes(x=(as.Date("2015-12-25")), y= 60, label='Christmas'),
            size=3, angle=90, vjust=-0.4, hjust=0)+geom_vline(aes(xintercept=(as.numeric(as.Date("2015-08-01")))), 
             linetype = 2, color = 'brown') + 
  geom_text(aes(x=(as.Date("2015-08-01")), y= 60, label='Fall'),
            size=3, angle=90, vjust=-0.4, hjust=0)+geom_vline(aes(xintercept=(as.numeric(as.Date("2015-05-01")))), 
             linetype = 2, color = 'brown') + 
  geom_text(aes(x=(as.Date("2015-05-01")), y= 60, label='Summer'),
            size=3, angle=90, vjust=-0.4, hjust=0)+geom_vline(aes(xintercept=(as.numeric(as.Date("2015-03-01")))), 
             linetype = 2, color = 'brown') + 
  geom_text(aes(x=(as.Date("2015-03-01")), y= 60, label='Spring'),
            size=3, angle=90, vjust=-0.4, hjust=0)+geom_vline(aes(xintercept=(as.numeric(as.Date("2015-11-15")))), 
             linetype = 2, color = 'brown') + 
  geom_text(aes(x=(as.Date("2015-11-15")), y= 60, label='Thanksgiving and Winter'),
            size=3, angle=90, vjust=-0.4, hjust=0)

```


```{r fig.width=14, fig.height=7}
#This cell takes time please be patient:
gg_dob = grid.arrange(plot1, plot2, nrow = 1)
gg_dob
```


#Plot 10:
#Behavioral patterns in patients count:

```{r fig.width=5,fig.height=7}
pl1 <- ggplot(no_show_df)+geom_bar(aes(Scholarship, fill = Scholarship))+ylab('Count')+xlab("Scholarship")+scale_x_discrete(limits = c(0,1),labels = c("No","Yes"))+scale_y_continuous(limits = c(0,300000),label = unit_format(unit = "K", scale =1e-3))
pl2 <- ggplot(no_show_df)+geom_bar(aes(HiperTension, fill =HiperTension))+
    ylab('Count')+xlab("HyperTension")+scale_x_discrete(limits = c(0,1),labels = c("No","Yes"))+scale_y_continuous(limits = c(0,300000),label = unit_format(unit = "K", scale =1e-3))
pl3 <- ggplot(no_show_df)+geom_bar(aes(Diabetes, fill = Diabetes))+
    ylab('Count')+xlab("Diabetes")+scale_x_discrete(limits = c(0,1),labels = c("No","Yes"))+scale_y_continuous(limits = c(0,300000),label = unit_format(unit = "K", scale =1e-3))
pl4 <- ggplot(no_show_df)+geom_bar(aes (Alcoolism, fill =  Alcoolism))+
    ylab('Count')+xlab("Alcoholism")+scale_x_discrete(limits = c(0,1),labels = c("No","Yes"))+scale_y_continuous(limits = c(0,300000),label = unit_format(unit = "K", scale =1e-3))
pl5 <- ggplot(no_show_df)+geom_bar(aes( Handcap, fill = Handcap))+
    ylab('Count')+xlab("Handicap")+scale_x_discrete(limits = c(0,1,2,3,4))+scale_y_continuous(limits = c(0,300000),label = unit_format(unit = "K", scale =1e-3))
pl6 <- ggplot(no_show_df)+geom_bar(aes(Sms_Reminder, fill = Sms_Reminder))+
    ylab('Count')+xlab("Sms_Reminder")+scale_y_continuous(limits = c(0,300000),label = unit_format(unit = "K", scale =1e-3))
pl7 <- ggplot(no_show_df)+geom_bar(aes(Smokes, fill = Smokes))+
    ylab('Count')+xlab("Smokes")+scale_x_discrete(limits = c(0,1),labels = c("No","Yes"))+scale_y_continuous(limits = c(0,300000),label = unit_format(unit = "K", scale =1e-3))
pl8 <- ggplot(no_show_df)+geom_bar(aes(Tuberculosis, fill = Tuberculosis))+
    ylab('Count')+xlab("Tuberculosis")+scale_x_discrete(limits = c(0,1),labels = c("No","Yes"))+scale_y_continuous(limits = c(0,300000),label = unit_format(unit = "K", scale =1e-3))

fig = ggarrange(pl1,pl2,pl3,pl4,pl5,pl6,pl7,pl8, nrow = 4,ncol =2)
gg_ptmarkers = annotate_figure(fig, top = text_grob("Number of Patients in each marker", color = "black", size = 13))
gg_ptmarkers
```

#Plot 11:
#Behavioral patterns in patients proportion:


```{r fig.width=5,fig.height=7}
#str(no_show_df)
p1 <- ggplot(no_show_df)+geom_bar(aes(Scholarship, fill = Status), position = position_fill())+
    ylab('Proportion')+scale_x_discrete(limits = c(0,1),labels = c("No","Yes"))+xlab("Scholarship")+theme(legend.position="none")
p2 <- ggplot(no_show_df)+geom_bar(aes(HiperTension, fill =Status), position = position_fill())+
    ylab('Proportion')+scale_x_discrete(limits = c(0,1),labels = c("No","Yes"))+xlab("HyperTension")+theme(legend.position="none")
p3 <- ggplot(no_show_df)+geom_bar(aes(Diabetes, fill = Status), position = position_fill())+
    ylab('Proportion')+scale_x_discrete(limits = c(0,1),labels = c("No","Yes"))+xlab("Diabetes")+theme(legend.position="none")
p4 <- ggplot(no_show_df)+geom_bar(aes (Alcoolism, fill =  Status), position = position_fill())+
    ylab('Proportion')+scale_x_discrete(limits = c(0,1),labels = c("No","Yes"))+xlab("Alcoholism")+theme(legend.position="none")
p5 <- ggplot(no_show_df)+geom_bar(aes( Handcap, fill = Status), position = position_fill())+
    ylab('Proportion')+xlab("Handicap")+theme(legend.position="none")
p6 <- ggplot(no_show_df)+geom_bar(aes(Sms_Reminder, fill = Status), position = position_fill())+
    ylab('Proportion')+xlab("Sms_Reminder")+theme(legend.position="none")
p7 <- ggplot(no_show_df)+geom_bar(aes( Smokes, fill = Status), position = position_fill())+
    ylab('Proportion')+scale_x_discrete(limits = c(0,1),labels = c("No","Yes"))+xlab("Smokes")+theme(legend.position="none")
p8 <- ggplot(no_show_df)+geom_bar(aes( Tuberculosis, fill = Status), position = position_fill())+
    ylab('Proportion')+scale_x_discrete(limits = c(0,1),labels = c("No","Yes"))+xlab("Tuberculosis")


#ggarrange(p1,p2,p3,p4,p5,p6,p7,p8, nrow = 4, ncol=2, common.legend = TRUE, legend="bottom")+ggtitle("Behavioral Patterns in Patients")


figure <- ggarrange(p1,p2,p3,p4,p5,p6,p7,p8, nrow = 4, ncol=2, common.legend = TRUE, legend="right")

# Annotate the figure by adding a common labels
gg_ptmarkers_prop = annotate_figure(figure, top = text_grob("Behavioral Patterns in Patients", color = "black", size = 13))
gg_ptmarkers_prop


```




#Plot 12 Imbalance in data:
#No show and show 
```{r}

gg_imbalance = ggplot(data = no_show_df)+
    geom_bar(aes(x = Status,fill = Status))+
    ggtitle("Imbalance in Status variable")
gg_imbalance

```


#Plot 13:
#Gender plot


```{r}

##Gender plot:

no_shows_by_gender <- group_by(no_show_df, Gender)
no_shows_by_gender <- dplyr::summarize(no_shows_by_gender,
                                      absence_rate = mean(Status_int),
                                      n = n())

gender_labels = c('Male', 'Female')
no_shows_by_gender$gender_name <- gender_labels
no_shows_by_gender$gender_name <- 
  factor(no_shows_by_gender$gender_name, 
         levels=unique(no_shows_by_gender$gender_name))

gg_gender = ggplot(no_shows_by_gender, aes(x = gender_name, y = 100*(absence_rate-1))) + 
  geom_bar(stat = "identity",width = 0.5,fill = "blue" )  + 
  labs(x = 'Gender', y = 'Absence Rate %', 
       title='Absence Rate vs Gender') 

gg_gender


gg_gen_imb = ggplot(data = no_show_df)+
    geom_bar(aes(x = Gender,fill = Gender))+
    ggtitle("Imbalance in gender variable")

gg_gen_imb
```



#----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------#

##Research Questionnaire:

## Q1: What factors are most likely to determine whether a patient shows up to their scheduled doctor's appointment?

```{r fig.width=5,fig.height=7}
gg_mat
gg_age
gg_wait
gg_month_line
gg_dow
gg_appt_day
gg_sms_prop
gg_age_sms
gg_dob

```


## Q2: How is the absence/no-show to a scheduled appointment dependent on the general characteristics and behavior patterns of the patient?


```{r fig.width=5,fig.height=7}

gg_ptmarkers
gg_ptmarkers_prop

```

## Q3: Can we predict whether a patient would show up or not by taking the aforementioned variables as explanatory variables into consideration?

```{r fig.width=5,fig.height=7}
gg_imbalance

```


## The third question has been entirely dealt in the "pred_eda_proj.Rmd" file

